---
title: "Optic Nerve Hets vs WT DEG list"
author: "Jonathan Nelson and Katie Emberley"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if(!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}

if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}

if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("ggupset")) {install.packages("ggupset"); require("ggupset")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}

if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

if (!require("NatParksPalettes")) {install.packages("NatParksPalettes"); require("NatParksPalettes")}

```


# Load Seurat Object and create lists
```{r load data, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

ON_Combined <- readRDS(here("Outputs","ON_Combined_Final.rds"))

ON_Combined
```

# Create Cluster List for Analysis
For any cluster and genotype that has less than 3 nuclei will throw an error in "FindMarkers" when determining DEGs between genotypes. Therefore, you must refine the list of clusters for the forloop to go through. In this instance we must remove "COP2" and "KOOLs" as there are less than 3 nuclei in the WT Cre group.
```{r}
# Create list of clusters for DEG analysis
ON_Combined_list <- unlist(levels(ON_Combined@meta.data$named.clusters))

# Identify clusters with less than 3 nuclei per cluster
md <- ON_Combined@meta.data %>% as.data.table
md[, .N, by = c("Cre", "named.clusters")] %>% kable()

# Remove clusters with less than 3 nuclei for a genotype
ON_Combined_list
ON_Combined_list <- ON_Combined_list[-c(3, 7)]
ON_Combined_list
```

# PLP Cre+ AND Sox10 Cre+ (Het) vs. Control (WT) DEGs
```{r echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

DefaultAssay(ON_Combined) <- "RNA"

#Create a new environment for DEG lists
DEG.env <- new.env()
 
ls(DEG.env)

names(ON_Combined_list) <- ON_Combined_list

for(i in 1:length(ON_Combined_list)){

#Creating DEG list
markers <- FindMarkers(ON_Combined, ident.1 = "Het" , ident.2 = "WT", group.by = "Cre", subset.ident = ON_Combined_list[i], logfc.threshold = 0.322) 

#Sorting DEG list based on avg_log2FC from high to low
markers = markers[order(markers[,"avg_log2FC"], decreasing = TRUE),]

#Add Gene name to data frame and filter for only significant genes
markers <- markers %>%
  rownames_to_column(var="gene") %>%
  filter(p_val_adj < 0.05)

#Write object into environment
assign(paste("Sox10Cre", names((ON_Combined_list)[i]), sep = "_"), markers, envir = DEG.env)

#Print out top 50 DEGs
print(paste(ON_Combined_list[i], "Top 50 DEGs"))
top50 <- head(markers, 50)

#Prints the list of top 50 DEGs
print(top50)

#Create FeaturePlot and VlnPlot for top3DEGs
top3 <- head(markers$gene, 3)

for (j in top3) {
  A <- FeaturePlot(ON_Combined, j, split.by = "Cre")+ labs(subtitle= names(ON_Combined_list)[i])
  B <- VlnPlot(ON_Combined, j, split.by = "Cre", idents = names(ON_Combined_list)[i] , cols = natparks.pals("CraterLake")) + labs(title= j) + xlab(element_blank()) 
  print(A) + 
  print(B)
  }

#Create FeaturePlot and VlnPlot for bottom3 DEGs
bottom3 <- tail(markers$gene, 3)

for (j in bottom3) {
  C <- FeaturePlot(ON_Combined, j, split.by = "Cre")+ labs(subtitle= names(ON_Combined_list)[i])
  D <- VlnPlot(ON_Combined, j, split.by = "Cre", idents = names(ON_Combined_list)[i] , cols = natparks.pals("CraterLake")) + labs(title= j) + xlab(element_blank())
  print(C) + 
  print(D)
  }
}

list_objects <- ls(DEG.env)
save(list = list_objects, file = here("Outputs", "DEGs", "ON Combined Het vs WT Defining DEGs.RData"), envir = DEG.env)
 
allobj <- ls(DEG.env)
alldf <- sapply(allobj, get, simplify = FALSE, envir = DEG.env)
alldf <- alldf[sapply(alldf, inherits, "data.frame")]
alldf <- alldf[order(as.numeric(names(alldf)))]
 
write.xlsx(alldf, sheetName = names(alldf), file = here("Outputs","DEGs", "ON Combined Het vs WT Defining DEGs.xlsx"))
```

## Session Info

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
sessionInfo()
```


  