---
title: "Optic Nerve Combined: Microglia DEG Pathway Analysis"
author: "Katie Emberley, Jonathan Nelson, and Jeremiah Reyes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Packages
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if(!require("SeuratData")){install.packages("SeuratData"); require("SeurateData")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("writexl")) {install.packages("writexl"); require("writexl")}

if (!require("AnnotationHub")) {BiocManager::install('AnnotationHub'); require("AnnotationHub")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("ggupset")) {install.packages("ggupset"); require("ggupset")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("org.Mm.eg.db")) {BiocManager::install('org.Mm.eg.db'); require("org.Mm.eg.db")}
if (!require("tibble")) {install.packages("tibble"); require("tibble")} # rownames_to_column

if(!require("ggforce")){install.packages("ggforce"); require("ggforce")}
if(!require("graphlayouts")){install.packages("graphlayouts"); require("graphlayouts")}

if (!require("clusterProfiler")) {install.packages("clusterProfiler"); require("clusterProfiler")} # bitr


if(!require("NatParksPalettes")){(install.packages("NatParksPalettes")); require("NatParksPalettes")}

# Required for lineage tracing
if (!require("SeuratWrappers")) {install.packages("SeuratWrappers"); require("SeuratWrappers")}
if (!require("Matrix")) {install.packages("Matrix"); require("Matrix")}
if (!require("Signac")) {install.packages("Signac"); require("Signac")}
if (!require("SeuratData")) {install.packages("SeuratData"); require("SeuratData")}
if (!require("monocle3")) {install.packages("monocle3"); require("monocle3")}

```

# Load Subclustered MG object
```{r}
ON_MG <- readRDS(here("Outputs","Subclusters", "Microglia_Subclustering.rds"))
```

# Gene Set Enrichment Analysis for Clusters
ClusterProfiler paper: G Yu, LG Wang, Y Han, QY He. clusterProfiler: an R package for comparing biological themes among gene clusters. OMICS: A Journal of Integrative Biology 2012, 16(5):284-287. doi:[10.1089/omi.2011.0118](http://dx.doi.org/10.1089/omi.2011.0118)

ClusterProfiler tutorial: https://bioc.ism.ac.jp/packages/3.7/bioc/vignettes/enrichplot/inst/doc/enrichplot.html

## Reload RData for Cluster Defining DEGs and Change Object Names
```{r}
load(file = here("Outputs","DEGs","Microglia Subclustering DEGs", "Subclustering Microglia Cluster Defining DEGs.RData"))
Cluster0<-`0`
Cluster1<-`1`
Cluster2<-`2`
Cluster3<-`3`
Cluster4<-`4`

rm(`0`,`1`,`2`,`3`,`4`)
```

## Crude Pathways Cluster 0
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width = 6, fig.height = 8}

head(Cluster0, n = 50)

ENTREZ_list <- bitr(geneID = Cluster0$SYMBOL,
                    fromType = "SYMBOL",
                    toType = "ENTREZID",
                    OrgDb = "org.Mm.eg.db"
                    )

markers <-  ENTREZ_list %>% inner_join(Cluster0, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
head(markers, n = 50)

# Upregulated Genes in Homeostatic MG (0)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(pos.ranks)

pos_go <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(pos_go, split="ONTOLOGY") + 
 facet_grid(ONTOLOGY~., scale="free") + 
 ggtitle("Cluster 0 MG Upregulated Gene Pathways") + 
 theme(plot.title = element_text(hjust = 0.5))


pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20) +
  ggtitle("Cluster 0 MG Upregulated Gene Pathways") + 
  theme(axis.text.y = element_text(size = 12))
```

## Crude Pathways for Cluster 1
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width = 6, fig.height = 8}

head(Cluster1, n = 50)

ENTREZ_list <- bitr(geneID = Cluster1$SYMBOL,
                    fromType = "SYMBOL",
                    toType = "ENTREZID",
                    OrgDb = "org.Mm.eg.db"
                    )

markers <-  ENTREZ_list %>% inner_join(Cluster1, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
head(markers, n = 50)

# Upregulated Genes in Disease-Associated Microglia (Cluster1)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(pos.ranks)

pos_go <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(pos_go, split="ONTOLOGY") + 
 facet_grid(ONTOLOGY~., scale="free") + 
 ggtitle("Cluster 1 MG Upregulated Defining Pathways") + 
 theme(plot.title = element_text(hjust = 0.5))



pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20) +
  ggtitle("CLuster 1 Upregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))


# Downregulated Genes in Homeostatic Microglia (Cluster 1)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(neg_go, split="ONTOLOGY") + 
  facet_grid(ONTOLOGY~., scale="free") + 
  ggtitle("CLuster 1 Downregulated Defining Pathways") + 
  theme(plot.title = element_text(hjust = 0.5))

neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)+
  ggtitle("Cluster 1 Downregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))

```

## Crude Pathways for Cluster 2
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width = 6, fig.height = 8}

head(Cluster2, n = 50)

ENTREZ_list <- bitr(geneID = Cluster2$SYMBOL,
                    fromType = "SYMBOL",
                    toType = "ENTREZID",
                    OrgDb = "org.Mm.eg.db"
                    )

markers <-  ENTREZ_list %>% inner_join(Cluster2, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
head(markers, n = 50)

# Upregulated Genes in Disease-Associated Microglia (Cluster2)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(pos.ranks)

pos_go <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(pos_go, split="ONTOLOGY") + 
 facet_grid(ONTOLOGY~., scale="free") + 
 ggtitle("Cluster 2 Upregulated Defining Pathways") + 
 theme(plot.title = element_text(hjust = 0.5))



pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20) +
  ggtitle("Cluster 2 Upregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))


# Downregulated Genes in Homeostatic Microglia (Cluster 1)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(neg_go, split="ONTOLOGY") + 
  facet_grid(ONTOLOGY~., scale="free") + 
  ggtitle("CLuster 2 Downregulated Defining Pathways") + 
  theme(plot.title = element_text(hjust = 0.5))

neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)+
  ggtitle("Cluster 2 Downregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))

```

## Crude Pathways for Cluster 3
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width = 6, fig.height = 8}

head(Cluster3, n = 50)

ENTREZ_list <- bitr(geneID = Cluster3$SYMBOL,
                    fromType = "SYMBOL",
                    toType = "ENTREZID",
                    OrgDb = "org.Mm.eg.db"
                    )

markers <-  ENTREZ_list %>% inner_join(Cluster3, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
head(markers, n = 50)

# Upregulated Genes in Disease-Associated Microglia (Cluster3)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(pos.ranks)

pos_go <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(pos_go, split="ONTOLOGY") + 
 facet_grid(ONTOLOGY~., scale="free") + 
 ggtitle("Cluster 3 Upregulated Defining Pathways") + 
 theme(plot.title = element_text(hjust = 0.5))



pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20) +
  ggtitle("Cluster 3 Upregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))


# Downregulated Genes in Homeostatic Microglia (Cluster 1)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(neg_go, split="ONTOLOGY") + 
  facet_grid(ONTOLOGY~., scale="free") + 
  ggtitle("CLuster 3 Downregulated Defining Pathways") + 
  theme(plot.title = element_text(hjust = 0.5))

neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)+
  ggtitle("Cluster 3 Downregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))

```

## Crude Pathways for Cluster 4
```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width = 6, fig.height = 8}

head(Cluster4, n = 50)

ENTREZ_list <- bitr(geneID = Cluster4$SYMBOL,
                    fromType = "SYMBOL",
                    toType = "ENTREZID",
                    OrgDb = "org.Mm.eg.db"
                    )

markers <-  ENTREZ_list %>% inner_join(Cluster4, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
head(markers, n = 50)

# Upregulated Genes in Disease-Associated Microglia (Cluster4)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(pos.ranks)

pos_go <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(pos_go, split="ONTOLOGY") + 
 facet_grid(ONTOLOGY~., scale="free") + 
 ggtitle("Cluster 4 Upregulated Defining Pathways") + 
 theme(plot.title = element_text(hjust = 0.5))



pos_go.BP <- enrichGO(pos.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(pos_go.BP, showCategory=20) +
  ggtitle("Cluster 4 Upregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))


# Downregulated Genes in Homeostatic Microglia (Cluster 1)

neg.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
head(neg.markers, n = 50)

neg.ranks <- neg.markers$ENTREZID[abs(markers$avg_log2FC) > 0.3219]
head(neg.ranks)

neg_go <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont="all",
                readable=TRUE)

dotplot(neg_go, split="ONTOLOGY") + 
  facet_grid(ONTOLOGY~., scale="free") + 
  ggtitle("CLuster 4 Downregulated Defining Pathways") + 
  theme(plot.title = element_text(hjust = 0.5))

neg_go.BP <- enrichGO(neg.ranks,
                OrgDb = "org.Mm.eg.db",
                ont ="BP",
                readable=TRUE)

barplot(neg_go.BP, showCategory=20)+
  ggtitle("Cluster 4 Downregulated Defining Pathways") + 
  theme(axis.text.y = element_text(size = 12))

```

# Session Info
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()
```

